---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(ggplot2)
library(dplyr)
library(data.table)
unzip('activity.zip')
data = read.csv('activity.csv', header = TRUE)
data$date = as.Date(data$date)

```

## What is mean total number of steps taken per day?

```{r}
totalsteps = tapply(data$steps, data$date, sum)
qplot(totalsteps, geom = 'histogram', xlab = 'number of steps', main = 'Steps Per Day', bins = 10)
mean = mean(totalsteps, na.rm = TRUE)
median = median(totalsteps, na.rm = TRUE)

```

This is the mean number of steps:

```{r}
mean
```

This is the median number of steps:

```{r}
median
```

## What is the average daily activity pattern?

```{r}
intervals = tapply(data$steps, data$interval, mean, na.rm = TRUE)
plot(as.numeric(names(intervals)),
     intervals, xlab = 'Interval', ylab = 'number of steps', main = 'Average Daily Activity Pattern')
maxinterval = names(sort(intervals, decreasing = TRUE))
maxnumber = sort(intervals, decreasing = TRUE)[1]
```
This is the 5-minute interval with the most # steps and the maximum value:

```{r}
maxnumber
```

## Imputing missing values

```{r}
missingvals = is.na(data$steps)
totalmissing = sum(missingvals)
```

This is the total number of missing values:
```{r}
totalmissing
```

My strategy is to fill in missing data with the average of the available data points for each interval

```{r}
datasplit = split(data, data$interval)
for(i in 1:length(datasplit)){
  datasplit[[i]]$steps[is.na(datasplit[[i]]$steps)] = intervals[i]
  
}

datainput = do.call('rbind', datasplit)
datainput = datainput[order(datainput$date) ,]

totalstepsinput = tapply(datainput$steps, datainput$date, sum)
qplot(totalstepsinput, geom = 'histogram', xlab = 'number of steps', main = 'Steps Per Day (Missing Data Filled In)', bins = 10)

mean = mean(totalstepsinput, na.rm = TRUE)
median = median(totalstepsinput, na.rm = TRUE)

```

This is the mean number of steps with the missing values filled in:
```{r}
mean
```

This is the median number of steps with the missing values filled in:
```{r}
median
```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
days = mutate(data, day = ifelse(weekdays(data$date) == 'Saturday' | weekdays(data$date) == 'Sunday', 'weekend', 'weekday'))
days$day = as.factor(days$day)

wknd = subset(days,as.character(days$day)=='weekend')
wkdy = subset(days,as.character(days$day)=='weekday')
wkndsteps = with(wknd,tapply(steps,interval,mean,na.rm=TRUE))
wkdysteps = with(wkdy,tapply(steps,interval,mean,na.rm=TRUE))
intwknd = unique(wknd$interval)
intwkdy = unique(wkdy$interval)
newwknd = data.frame(cbind(wkndsteps,intwknd))
newwkdy = data.frame(cbind(wkdysteps,intwkdy))
par(mfrow=c(2,1),mar=c(4,4,2,1))
plot(newwknd$intwknd,newwknd$wkndsteps,xlab = 'intervals', ylab = 'steps mean', main = 'weekend')
plot(newwkdy$intwkdy,newwkdy$wkdysteps,xlab = 'intervals', ylab = 'steps mean', main = 'weekday')
```
